---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data (i.e. read.csv())
```{r}
if(!exists("./activity.zip")){
    fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(fileUrl, destfile = "activity.zip", method = "curl")
}

activityUnzip <- unzip(("activity.zip"), files = "activity.csv")

activityData <- read.csv(activityUnzip, sep = ",", header = TRUE)
```

2. Process/transform the data (if necessary) into a format suitable for your analysis

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r}
activityDataDailySteps <- aggregate(cbind(steps) ~ date, data = activityData, FUN = sum)
```

2. Make a histogram of the total number of steps taken each day

```{r}
barplot(activityDataDailySteps$steps,
        names.arg = activityDataDailySteps$date,
        xlab = "Date",
        ylab = "Steps",
        main = "Total Number of Steps Taken Each Day")
```

3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
activityDataDailySteps
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
